Distributed Nonlinear Feedback Design with Separable Control-Contraction Metrics

### 2016-11-21 R0V0.tex
Add paragraph numbering

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% 																			   %
% Git Branch: Manchester20161020											   %
% 																			   %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

@@

Over the years, methods for the design of controllers for dynamical systems has developed different techniques, according to the vector field that describes the system structure. The main underlying concept employed for the synthesis of feedback laws is the Lyapunov's stability theory \cite{Bhatia1970,Rouche:1977}. In terms of computation, controller design problems are often translated in terms of dynamic programming.

For linear systems, controller design methods include $H_\infty$ optmization \cite{Doyleetal1992} and LQR \cite{Hespanha:2009}. The design problems are usually formulated in terms of linear matrix inequalities (LMIs) and semidefinite programming \cite{Boyd1994,DullerudPaganini2000}. In the case of nonlinear systems, design approaches include high-gain  \cite{GrognardSepulchreBastin1999}, backstepping \cite{Krstic1995}, and forwarding \cite{MazencPraly96}. These methods depend on the structure of the vector field and may fail to be applied when appropriate conditions are not met \cite{SteinShiromoto2013a}. In terms of numerical computation, often the techniques employed for feedback synthesis include sum-of-squares (SOS) optimization \cite{Parrilo2003}.

For those systems described by input-affine differential equations,  the existence of a control-Lyapunov function (CLF) satisfying the so-called \emph{small control property} provides a ``universal'' formula for a stabilizing feedback law  \cite{Sontag1983,Sontag1989a}. However, by the time that this paper was submitted, systematic methods for the search of a CLF were known only for particular systems within this class (see, for instance, \cite{MazencMalisoff2006}). In addition to the lack of a general systematic approach, the search for a CLF in terms of numerical optimization leads to a problem that is not necessarily convex nor connected \cite{Rantzer:2001}.

An alternative to CLFs is provided by the so-called control-contraction metrics (CCMs). One of the main advantages of CCMs over CLFs is that the feedback law is obtained by solving a convex optimization problem \cite{Manchester2014a}, due to fact that the \emph{differential / variational} formulation (see definition below) of the system is considered. The obtained controller makes every solution of the closed-loop system exponentially converge to a prescribed solution (see a precise definition below). In contrast to feedback linearization which depends on integrability, the obtained controller depends only on the stabilizability property of the system. This stability notion is related to contraction theory (an incremental stability property \cite{Angeli2002,Lohmiller1998,Rueffer2013,Sontag2010,Forni2014}) which can be traced back to the work \cite{Lewis1949}.

The drawback of methods such as CCMs, $H_\infty$ optimization and LQR is that, when the system under consideration has many states (large scale), it does not scale well. This is due to the fact that the stability verification of a system with $n$ states may require a function with $n^2$ terms (see \cite{Rantzer2015} for the case of linear systems and Lyapunov functions).

Large-scale systems may appear in the form of interconnection of smaller components, for example, in applications such as traffic networks \cite{CanudasdeWitMorbidiLeonOjedaEtAl2015}, pipe networks \cite{Persis2011a} and sensor networks \cite{Pajic2011}. For such systems, a property often desired is the ability to synthetize controllers with particular structures (see \cite{Siljak1991,Bai2011} for distributed controllers).

In terms of dynamic programming, the problem of design controllers having an arbitrary structure is known to be NP-hard \cite{BlondelTsitsiklis1997,Tanaka2011}, for linear systems. Consequently, for nonlinear systems the problem is very challenging.

In the case of linear systems whose solutions evolve in the positive orthant (positive systems), the problem of design a structured stabilizing feedback law has been addressed in \cite{Tanaka2011}. Also, the authors show that the existence of such a controller is equivalent to the existence of Lyapunov function with a diagonal structure. The importance of this result for large-scale systems is highlighted in \cite{Rantzer2015}, where sparsity is explored to reduce the number of LMIs to be solved.

For nonlinear systems described by a non-fully connected graph, the design of structured controllers could be simplified by choosing structured Lyapunov functions (see \cite{Dirr2015} and references therein). However, the search for a general Lyapunov function is a non-trivial problem by itself \cite{MalisoffMazenc2009}. The case can be even worse, when structural constraints are considered.

For this reason, the structured design of feedback laws for nonlinear systems is investigated with CCMs satisfying structural constraints, in this work. The systems under consideration are continuously differentiable and affine in the input. One of the key elements employed to design these controllers is the concept of a Riemannian metric (see the definition below) which is employed, similarly to Lyapunov functions, to measure the exponential convergence of pairs of solutions. Employing CCMs that can be decomposed into the sum of smaller elements, the design of a structured feedback law allowing exponential convergence of solutions is obtained in two steps.

The first step consists of an off-line computation. The original system is extended with its \emph{differential / variational} formulation. The exponential convergence problem is reformulated in terms of stability of the origin for the extended system. This stability problem is written in terms of a convex optimization problem under the prescribed structural constraint. The solution consists of a pair of matrices: the Riemannian metric and the \emph{differential / variational} feedback law. The second step consists of an on-line optimization. The obtained \emph{differential / variational} controller is integrated along geodesics (see definition below).

At this point, the three contributions of this work can be highlighted. First, using the methodology for synthesis of CCMs introduced in \cite{Manchester2014a} and imposing a (block-)diagonal structure constraint, a feedback law with prescribed structure is obtained for nonlinear systems. Second, because the CCM has a (block-)diagonal structure, the computation of the online integration phase, which corresponds to an optimization problem, is distributed. Third, when the system is described a non-strongly connected graph, the structure imposed on the CCM allows to reduce the number of equations needed to  verify stability. Note that the proposed method provides an explicit algorithm to design the controller which makes this approach constructive.

Other works related to analysis under structural constraints to achieve exponential convergence of solutions do exist in the literature. For contraction analysis, see \cite{Aminzare2014a} for analysis according to different network topologies and \cite{Andrieu2016} for transversal contraction analysis. This paper contrast with these works mainly in two points: 1) A  methodology for the design of structured controllers is proposed; 2) Due the formulation of the stabilization problem as a convex optimization problem, an explicit formula of a structured controller is obtained.

@@

The notation $\mathcal{L}_{\mathrm{loc}}^\infty(\mathbb{R}_{\geq0},\mathbb{R}^m)$ stands for the class of functions $u:\mathbb{R}\to\mathbb{R}^m$ that are locally essentially bounded. Given differentiable functions $M:\mathbb{R}^n\to\mathbb{R}^{n\times n}$ and $f:\mathbb{R}^n\to\mathbb{R}^n$ the notation $\partial_fM$ stands for matrix with dimension $n\times n$ and with $(i,j)$ element given by $\frac{\partial m_{ij}}{\partial x}(x)f(x)$. The notation $\dot{f}$ stands for the total derivative of $f$.

Let $N>0$ be an integer, a \emph{graph} consists of a set of \emph{nodes} $\mathscr{V}\subset\mathbb{N}_{[1,N]}$ and a set of \emph{edges} $\mathscr{E}\subset\mathscr{V}\times\mathscr{V}$ and it is denoted by the pair $(\mathscr{V},\mathscr{E})=\mathscr{G}$. A node $i\in\mathscr{V}$ is said to be \emph{adjacent} to a node $j\in\mathscr{V}$ if $(i,j)\in\mathscr{E}$, the set of nodes that are adjacent to $j$ is defined as $\mathscr{N}(j)=\{i\in\mathscr{V}:i\neq j,(i,j)\in\mathscr{E}\}$. Given two nodes $i,j\in\mathscr{V}$, and ordered sequence of edges $\left\{(k,k+1)\right\}_{k=i}^{j-1}$ is said to be a \emph{path from the node $i$ to the node $j$}. A path is said to be \emph{cycle} if no edges are repeated and the nodes $i$ and $j-1$  are distinct. A graph is said to be \emph{strongly connected} if, for every two nodes $i,j\in\mathscr{V}$, there exists a path connecting them. It is also said to be \emph{complete} is every node is adjacent to any other node, and it is \emph{incomplete} otherwise. A graph is said to be a \emph{tree} if it is connected and does not contain cycles. A \emph{leaf} is a node that is adjacent to only one node.

@@

Consider the class of systems described by the differential equation
\begin{equation}\label{eq:general system}
	\dot{x}(t)=f(x(t))+B(x(t))u(t)\;,	
\end{equation}
where, for every positive times $t$, the \emph{system state} $x(t)$ and the \emph{system input variable} $u(t)$ evolve in the Euclidean spaces $\mathbb{R}^n$ and $\mathbb{R}^m$, respectively. The functions $f:\mathbb{R}^n\to\mathbb{R}^n$ and $B:\mathbb{R}^n\to\mathbb{R}^m$ are assumed to be smooth, i.e., infinitely differentiable. From now on the dependence on the time $t$ will be omitted.

A function $u^\ast\in\mathcal{L}_{\mathrm{loc}}^\infty(\mathbb{R}_{\geq0},\mathbb{R}^m)$ is said to be an \emph{input signal or control for \eqref{eq:general system}}. For such a control for \eqref{eq:general system}, and for every \emph{initial condition} $x^\ast$, there exists a unique solution to \eqref{eq:general system} (\cite{Teschl2012}) that is denoted by $X(t,x^\ast,u^\ast)$, when computed at time $t$. This solution is defined over an open interval $(\underline{t},\overline{t})$, and it is said to be \emph{forward complete} if $\overline{t}=+\infty$.

A forward complete solution $X^\ast(\cdot,x^\ast,u^\ast)$ to \eqref{eq:general system} is said to be  \emph{globally exponentially uniformly stabilizable} with \emph{rate} $\lambda>0$ if there exist a constant value $C>0$ and a feedback law $k^\ast:\mathbb{R}_{\geq0}\times\mathbb{R}^n\times\{X^\ast\}\times\{u^\ast\}\to\mathbb{R}^m$, denoted as $k^\ast$, such that the inequality
\begin{equation}\label{eq:global exponential uniform stabilizability}
	\left|X^\ast(t,x^\ast,u^\ast)-X(t,x,k^\ast)\right|\leq Ce^{-\lambda t}|x^\ast-x|
\end{equation}
holds, for every $t\geq0$, and for every initial condition $x\in\mathbb{R}^n$ (\cite{Manchester2014a}). Equivalently, if inequality~\eqref{eq:global exponential uniform stabilizability} holds, then the solution $X^\ast(\cdot,x^\ast,u^\ast)$ to \eqref{eq:general system} is said to be \emph{globally exponentially uniformly stable} for system \eqref{eq:general system} in closed loop with $k^\ast$

A stronger condition than the global exponential stabilizability of a particular solution is the requirement for every forward complete solution to be globally exponentially stabilizable. This concept is formalized in the following definition recalled from \cite{Manchester2014a}.

@@

Let $N>0$ be an integer, for each index $i\in\mathbb{N}_{[1,N]}$, system~\eqref{eq:general system} can be decomposed into smaller components as follows.
\begin{equation}\label{eq:subsystem}
	\dot{x}_i=f_i(x_i,\breve{x}_i)+b_i(x_i)u_i\;,
\end{equation}
where $\breve{x}_i\in\mathbb{R}^{\breve{n}_i}$ stands for the vector composed of states $x_j$ of systems that are adjacent to system~\eqref{eq:subsystem}.

At this point, the problem under consideration in this paper can be stated as follows.

@@

The constraint on the components of $k^\ast$ described in Problem~\ref{problem formulation} is particularly relevant for the design feedback laws with a prescribed structure (topology). Consider the network composed of systems described by the following equation.
\begin{equation}\label{eq:example}
	\scalebox{0.95}{$\left\{\begin{array}{rcl}
		\dot{x}_i&=&-x_i-x_i^3+y_i^2 + 0.01\left(x_{i-1}^3 - 2x_i^3 +x_{i+1}^3\right)\\
		\dot{y}_i&=&u_i\;.
	\end{array}\right.$}
\end{equation}
As explained in Section \ref{sec:Illustration}, when more than five of these systems are connected, the CCM approach was unable to provide unconstrained controller on a standard desktop machine, due to the number of variables. However, by exploring sparsity, up to 512 of theses systems can be connected and the controller can be computed.

@@

A \emph{Riemannian metric on $\mathbb{R}^n$} is a symmetric positive-definite bilinear form that depends smoothly on $x\in\mathbb{R}^n$. In a particular coordinate system, for any pair of vectors $\delta_0,\delta_1$ of $\mathbb{R}^n$ the \emph{metric} is defined as the inner product $\langle\delta_0,\delta_{ 1}\rangle_x=\delta_0^\top M(x)\delta_1$, where $M:\mathbb{R}^n\to\mathbb{R}^{n\times n}$ is a smooth function. Consequently, local notions of norm $|\delta_0|_x^2=\langle\delta_0,\delta_0\rangle_x=:V(x,\delta_0)$ and orthogonality $\langle\delta_0,\delta_1\rangle_x=0$ can be defined. The metric is said to be \emph{bounded} if there exists constant values $\underline{m}>0$ and $\overline{m}>0$ such that, for every $x\in\mathbb{R}^n$, $\underline{m}I_n\leq M(x)\leq \overline{m}I_n$, where $I_n\in\mathbb{R}^{n\times n}$ is the identity matrix. From now, metric stands for a bounded Riemannian metric.

Let $\Gamma(x_0,x_1)$ be the set of piecewise-smooth curves $c:[0,1]\to\mathbb{R}^n$ connecting $x_0=c(0)$ to $x_1=c(1)$. The \emph{length} and \emph{energy} of $c$ are, respectively, defined by the values
\begin{align*}
	\ell(\gamma)=&\ \int_0^1\left|c_s(s)\right|_{c(s)}\,ds\\
	e(\gamma)=&\ \int_0^1V\left(c(s),c_s(s)\right)\,ds\;,
\end{align*}
where the notation $c_s$ stands for the derivative $\tfrac{\partial c}{\partial s}$. The Riemannian distance between $x_0$ and $x_1$, denoted as $\mathbin{\mathtt{dist}}(x_0,x_1)$, is defined as the curve with the smallest length connecting them. This curve is said to be a \emph{geodesic} and it is the solution to the optimization problem
\begin{equation}\label{eq:geodesic formulation}
	\mathbin{\mathtt{dist}}(x_0,x_1)=\inf_{c\in\Gamma(x_0,x_1)}\ell(c)\;.
\end{equation}

A suitable notion to deal with exponential convergence of pair of solutions to \eqref{eq:general system} is provided by the \emph{differential} (also known as variational or prolonged) dynamical system
\begin{equation}\label{eq:general system:differential}
	\dot{\delta}_x=A(x,u)\delta_x+B(x)\delta_u\;,
\end{equation}
where $\delta_x$ (resp. $\delta_u$) is a vector of the Euclidean space $\mathbb{R}^n$ (resp. $\mathbb{R}^m$). More precisely, it is the vector tangent to a piecewise smooth curve connecting a pair of points in $\mathbb{R}^n$ (resp. $\mathbb{R}^m$). The matrix $A\in\mathbb{R}^{n\times n}$ has components given, for every $(x,u)\in\mathbb{R}^n\times\mathbb{R}^m$, by 
\begin{equation*}
	A_{jk}(x,u)=\dfrac{\partial[f_j+b_ju_j]}{\partial x_k}(x,u)	
\end{equation*}
for indexes $j,k\in\mathbb{N}_{[1,n]}$.

The resulting system composed of Equations \eqref{eq:general system} and \eqref{eq:general system:differential} is analyzed on the state space spanned by the vector $(x,\delta_x)\in\mathbb{R}^n\times\mathbb{R}^n$.

Similarly to \eqref{eq:general system}, given a control $\delta_u$ for system \eqref{eq:general system:differential}, the solution to \eqref{eq:general system:differential} computed at time $t\geq0$, $(x,u)\in\mathbb{R}^n\times\mathbb{R}^n$ and issuing from the  initial condition $\delta_x\in\mathbb{R}^n$ is denoted by $\Delta_x(t,x,\delta_x,u,\delta_u)$.

Lyapunov stability notions of solutions to \eqref{eq:general system:differential} are similar to those of linear time-varying systems (LTVS) (see \cite{Hespanha:2009} for more information on LTVS).

The importance of the stability of \eqref{eq:general system:differential} for system \eqref{eq:general system} can be understood as follows. Given fixed controls $u$ and $\delta u$ for systems \eqref{eq:general system} and \eqref{eq:general system:differential}, respectively. If the norm of every solution $\Delta_x(t,x,\delta_x,u,\delta_u)$ converges to zero exponentially, as the time tends to infinity, then every pair of solutions to \eqref{eq:general system} converge to each other exponentially. The interested reader may address \cite{Lohmiller1998,Sontag2010} and references therein for further details.

A sufficient condition for the stability of \eqref{eq:general system:differential} is provided by analyzing the derivative of a particular function along the solutions of systems \eqref{eq:general system} and \eqref{eq:general system:differential}. This function is recalled from \cite{Forni2014} and \cite{Manchester2014a}.

@@

The existence a contraction metric for system \eqref{eq:general system} implies that every two solutions of this system converge to each other exponentially. The proof of this claim, for the case in which $u\equiv0$ and $\delta_u\equiv0$, can be found in \cite[Theorem 1]{Lewis1951}, and \cite[Theorems 5.7 and 5.33]{Reich2005}, and \cite[Lemma 3.3]{Isac2008}.

For the class of systems considered in this paper, the following kind of metric is of interest, since it also allows the design a feedback law for system \eqref{eq:general system:differential}.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% 																			   %
% Git Branch: master														   %
% 																			   %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

% Defines a paragraph counter ----------------------------------------------

\usepackage{chngcntr} % To reset counter at each section

% Paragraph counter 
\newcounter{para}
\newcommand\mypara{\par \thesection.\refstepcounter{para}\thepara.\space}

\counterwithin*{para}{section} % Resets counter at each section

@@

\mypara Over the years, methods for the design of controllers for dynamical systems has developed different techniques, according to the vector field that describes the system structure. The main underlying concept employed for the synthesis of feedback laws is the Lyapunov's stability theory \cite{Bhatia1970,Rouche:1977}. In terms of computation, controller design problems are often translated in terms of dynamic programming.

\mypara For linear systems, controller design methods include $H_\infty$ optmization \cite{Doyleetal1992} and LQR \cite{Hespanha:2009}. The design problems are usually formulated in terms of linear matrix inequalities (LMIs) and semidefinite programming \cite{Boyd1994,DullerudPaganini2000}. In the case of nonlinear systems, design approaches include high-gain  \cite{GrognardSepulchreBastin1999}, backstepping \cite{Krstic1995}, and forwarding \cite{MazencPraly96}. These methods depend on the structure of the vector field and may fail to be applied when appropriate conditions are not met \cite{SteinShiromoto2013a}. In terms of numerical computation, often the techniques employed for feedback synthesis include sum-of-squares (SOS) optimization \cite{Parrilo2003}.

\mypara For those systems described by input-affine differential equations,  the existence of a control-Lyapunov function (CLF) satisfying the so-called \emph{small control property} provides a ``universal'' formula for a stabilizing feedback law  \cite{Sontag1983,Sontag1989a}. However, by the time that this paper was submitted, systematic methods for the search of a CLF were known only for particular systems within this class (see, for instance, \cite{MazencMalisoff2006}). In addition to the lack of a general systematic approach, the search for a CLF in terms of numerical optimization leads to a problem that is not necessarily convex nor connected \cite{Rantzer:2001}.

\mypara An alternative to CLFs is provided by the so-called control-contraction metrics (CCMs). One of the main advantages of CCMs over CLFs is that the feedback law is obtained by solving a convex optimization problem \cite{Manchester2014a}, due to fact that the \emph{differential / variational} formulation (see definition below) of the system is considered. The obtained controller makes every solution of the closed-loop system exponentially converge to a prescribed solution (see a precise definition below). In contrast to feedback linearization which depends on integrability, the obtained controller depends only on the stabilizability property of the system. This stability notion is related to contraction theory (an incremental stability property \cite{Angeli2002,Lohmiller1998,Rueffer2013,Sontag2010,Forni2014}) which can be traced back to the work \cite{Lewis1949}.

\mypara The drawback of methods such as CCMs, $H_\infty$ optimization and LQR is that, when the system under consideration has many states (large scale), it does not scale well. This is due to the fact that the stability verification of a system with $n$ states may require a function with $n^2$ terms (see \cite{Rantzer2015} for the case of linear systems and Lyapunov functions).

\mypara Large-scale systems may appear in the form of interconnection of smaller components, for example, in applications such as traffic networks \cite{CanudasdeWitMorbidiLeonOjedaEtAl2015}, pipe networks \cite{Persis2011a} and sensor networks \cite{Pajic2011}. For such systems, a property often desired is the ability to synthetize controllers with particular structures (see \cite{Siljak1991,Bai2011} for distributed controllers).

\mypara In terms of dynamic programming, the problem of design controllers having an arbitrary structure is known to be NP-hard \cite{BlondelTsitsiklis1997,Tanaka2011}, for linear systems. Consequently, for nonlinear systems the problem is very challenging.

\mypara In the case of linear systems whose solutions evolve in the positive orthant (positive systems), the problem of design a structured stabilizing feedback law has been addressed in \cite{Tanaka2011}. Also, the authors show that the existence of such a controller is equivalent to the existence of Lyapunov function with a diagonal structure. The importance of this result for large-scale systems is highlighted in \cite{Rantzer2015}, where sparsity is explored to reduce the number of LMIs to be solved.

\mypara For nonlinear systems described by a non-fully connected graph, the design of structured controllers could be simplified by choosing structured Lyapunov functions (see \cite{Dirr2015} and references therein). However, the search for a general Lyapunov function is a non-trivial problem by itself \cite{MalisoffMazenc2009}. The case can be even worse, when structural constraints are considered.

\mypara For this reason, the structured design of feedback laws for nonlinear systems is investigated with CCMs satisfying structural constraints, in this work. The systems under consideration are continuously differentiable and affine in the input. One of the key elements employed to design these controllers is the concept of a Riemannian metric (see the definition below) which is employed, similarly to Lyapunov functions, to measure the exponential convergence of pairs of solutions. Employing CCMs that can be decomposed into the sum of smaller elements, the design of a structured feedback law allowing exponential convergence of solutions is obtained in two steps.

\mypara The first step consists of an off-line computation. The original system is extended with its \emph{differential / variational} formulation. The exponential convergence problem is reformulated in terms of stability of the origin for the extended system. This stability problem is written in terms of a convex optimization problem under the prescribed structural constraint. The solution consists of a pair of matrices: the Riemannian metric and the \emph{differential / variational} feedback law. The second step consists of an on-line optimization. The obtained \emph{differential / variational} controller is integrated along geodesics (see definition below).

\mypara At this point, the three contributions of this work can be highlighted. First, using the methodology for synthesis of CCMs introduced in \cite{Manchester2014a} and imposing a (block-)diagonal structure constraint, a feedback law with prescribed structure is obtained for nonlinear systems. Second, because the CCM has a (block-)diagonal structure, the computation of the online integration phase, which corresponds to an optimization problem, is distributed. Third, when the system is described a non-strongly connected graph, the structure imposed on the CCM allows to reduce the number of equations needed to  verify stability. Note that the proposed method provides an explicit algorithm to design the controller which makes this approach constructive.

\mypara Other works related to analysis under structural constraints to achieve exponential convergence of solutions do exist in the literature. For contraction analysis, see \cite{Aminzare2014a} for analysis according to different network topologies and \cite{Andrieu2016} for transversal contraction analysis. This paper contrast with these works mainly in two points: 1) A  methodology for the design of structured controllers is proposed; 2) Due the formulation of the stabilization problem as a convex optimization problem, an explicit formula of a structured controller is obtained.

@@

\mypara The notation $\mathcal{L}_{\mathrm{loc}}^\infty(\mathbb{R}_{\geq0},\mathbb{R}^m)$ stands for the class of functions $u:\mathbb{R}\to\mathbb{R}^m$ that are locally essentially bounded. Given differentiable functions $M:\mathbb{R}^n\to\mathbb{R}^{n\times n}$ and $f:\mathbb{R}^n\to\mathbb{R}^n$ the notation $\partial_fM$ stands for matrix with dimension $n\times n$ and with $(i,j)$ element given by $\frac{\partial m_{ij}}{\partial x}(x)f(x)$. The notation $\dot{f}$ stands for the total derivative of $f$.

\mypara Let $N>0$ be an integer, a \emph{graph} consists of a set of \emph{nodes} $\mathscr{V}\subset\mathbb{N}_{[1,N]}$ and a set of \emph{edges} $\mathscr{E}\subset\mathscr{V}\times\mathscr{V}$ and it is denoted by the pair $(\mathscr{V},\mathscr{E})=\mathscr{G}$. A node $i\in\mathscr{V}$ is said to be \emph{adjacent} to a node $j\in\mathscr{V}$ if $(i,j)\in\mathscr{E}$, the set of nodes that are adjacent to $j$ is defined as $\mathscr{N}(j)=\{i\in\mathscr{V}:i\neq j,(i,j)\in\mathscr{E}\}$. Given two nodes $i,j\in\mathscr{V}$, and ordered sequence of edges $\left\{(k,k+1)\right\}_{k=i}^{j-1}$ is said to be a \emph{path from the node $i$ to the node $j$}. A path is said to be \emph{cycle} if no edges are repeated and the nodes $i$ and $j-1$  are distinct. A graph is said to be \emph{strongly connected} if, for every two nodes $i,j\in\mathscr{V}$, there exists a path connecting them. It is also said to be \emph{complete} is every node is adjacent to any other node, and it is \emph{incomplete} otherwise. A graph is said to be a \emph{tree} if it is connected and does not contain cycles. A \emph{leaf} is a node that is adjacent to only one node.

@@

\mypara Consider the class of systems described by the differential equation
\begin{equation}\label{eq:general system}
	\dot{x}(t)=f(x(t))+B(x(t))u(t)\;,	
\end{equation}
where, for every positive times $t$, the \emph{system state} $x(t)$ and the \emph{system input variable} $u(t)$ evolve in the Euclidean spaces $\mathbb{R}^n$ and $\mathbb{R}^m$, respectively. The functions $f:\mathbb{R}^n\to\mathbb{R}^n$ and $B:\mathbb{R}^n\to\mathbb{R}^m$ are assumed to be smooth, i.e., infinitely differentiable. From now on the dependence on the time $t$ will be omitted.

\mypara A function $u^\ast\in\mathcal{L}_{\mathrm{loc}}^\infty(\mathbb{R}_{\geq0},\mathbb{R}^m)$ is said to be an \emph{input signal or control for \eqref{eq:general system}}. For such a control for \eqref{eq:general system}, and for every \emph{initial condition} $x^\ast$, there exists a unique solution to \eqref{eq:general system} (\cite{Teschl2012}) that is denoted by $X(t,x^\ast,u^\ast)$, when computed at time $t$. This solution is defined over an open interval $(\underline{t},\overline{t})$, and it is said to be \emph{forward complete} if $\overline{t}=+\infty$.

\mypara A forward complete solution $X^\ast(\cdot,x^\ast,u^\ast)$ to \eqref{eq:general system} is said to be  \emph{globally exponentially uniformly stabilizable} with \emph{rate} $\lambda>0$ if there exist a constant value $C>0$ and a feedback law $k^\ast:\mathbb{R}_{\geq0}\times\mathbb{R}^n\times\{X^\ast\}\times\{u^\ast\}\to\mathbb{R}^m$, denoted as $k^\ast$, such that the inequality
\begin{equation}\label{eq:global exponential uniform stabilizability}
	\left|X^\ast(t,x^\ast,u^\ast)-X(t,x,k^\ast)\right|\leq Ce^{-\lambda t}|x^\ast-x|
\end{equation}
holds, for every $t\geq0$, and for every initial condition $x\in\mathbb{R}^n$ (\cite{Manchester2014a}). Equivalently, if inequality~\eqref{eq:global exponential uniform stabilizability} holds, then the solution $X^\ast(\cdot,x^\ast,u^\ast)$ to \eqref{eq:general system} is said to be \emph{globally exponentially uniformly stable} for system \eqref{eq:general system} in closed loop with $k^\ast$

\mypara A stronger condition than the global exponential stabilizability of a particular solution is the requirement for every forward complete solution to be globally exponentially stabilizable. This concept is formalized in the following definition recalled from \cite{Manchester2014a}.

@@

\mypara Let $N>0$ be an integer, for each index $i\in\mathbb{N}_{[1,N]}$, system~\eqref{eq:general system} can be decomposed into smaller components as follows.
\begin{equation}\label{eq:subsystem}
	\dot{x}_i=f_i(x_i,\breve{x}_i)+b_i(x_i)u_i\;,
\end{equation}
where $\breve{x}_i\in\mathbb{R}^{\breve{n}_i}$ stands for the vector composed of states $x_j$ of systems that are adjacent to system~\eqref{eq:subsystem}.

\mypara At this point, the problem under consideration in this paper can be stated as follows.

@@

\mypara The constraint on the components of $k^\ast$ described in Problem~\ref{problem formulation} is particularly relevant for the design feedback laws with a prescribed structure (topology). Consider the network composed of systems described by the following equation.
\begin{equation}\label{eq:example}
	\scalebox{0.95}{$\left\{\begin{array}{rcl}
		\dot{x}_i&=&-x_i-x_i^3+y_i^2 + 0.01\left(x_{i-1}^3 - 2x_i^3 +x_{i+1}^3\right)\\
		\dot{y}_i&=&u_i\;.
	\end{array}\right.$}
\end{equation}
As explained in Section \ref{sec:Illustration}, when more than five of these systems are connected, the CCM approach was unable to provide unconstrained controller on a standard desktop machine, due to the number of variables. However, by exploring sparsity, up to 512 of theses systems can be connected and the controller can be computed.

@@

\mypara A \emph{Riemannian metric on $\mathbb{R}^n$} is a symmetric positive-definite bilinear form that depends smoothly on $x\in\mathbb{R}^n$. In a particular coordinate system, for any pair of vectors $\delta_0,\delta_1$ of $\mathbb{R}^n$ the \emph{metric} is defined as the inner product $\langle\delta_0,\delta_{ 1}\rangle_x=\delta_0^\top M(x)\delta_1$, where $M:\mathbb{R}^n\to\mathbb{R}^{n\times n}$ is a smooth function. Consequently, local notions of norm $|\delta_0|_x^2=\langle\delta_0,\delta_0\rangle_x=:V(x,\delta_0)$ and orthogonality $\langle\delta_0,\delta_1\rangle_x=0$ can be defined. The metric is said to be \emph{bounded} if there exists constant values $\underline{m}>0$ and $\overline{m}>0$ such that, for every $x\in\mathbb{R}^n$, $\underline{m}I_n\leq M(x)\leq \overline{m}I_n$, where $I_n\in\mathbb{R}^{n\times n}$ is the identity matrix. From now, metric stands for a bounded Riemannian metric.

\mypara Let $\Gamma(x_0,x_1)$ be the set of piecewise-smooth curves $c:[0,1]\to\mathbb{R}^n$ connecting $x_0=c(0)$ to $x_1=c(1)$. The \emph{length} and \emph{energy} of $c$ are, respectively, defined by the values
\begin{align*}
	\ell(\gamma)=&\ \int_0^1\left|c_s(s)\right|_{c(s)}\,ds\\
	e(\gamma)=&\ \int_0^1V\left(c(s),c_s(s)\right)\,ds\;,
\end{align*}
where the notation $c_s$ stands for the derivative $\tfrac{\partial c}{\partial s}$. The Riemannian distance between $x_0$ and $x_1$, denoted as $\mathbin{\mathtt{dist}}(x_0,x_1)$, is defined as the curve with the smallest length connecting them. This curve is said to be a \emph{geodesic} and it is the solution to the optimization problem
\begin{equation}\label{eq:geodesic formulation}
	\mathbin{\mathtt{dist}}(x_0,x_1)=\inf_{c\in\Gamma(x_0,x_1)}\ell(c)\;.
\end{equation}

\mypara A suitable notion to deal with exponential convergence of pair of solutions to \eqref{eq:general system} is provided by the \emph{differential} (also known as variational or prolonged) dynamical system
\begin{equation}\label{eq:general system:differential}
	\dot{\delta}_x=A(x,u)\delta_x+B(x)\delta_u\;,
\end{equation}
where $\delta_x$ (resp. $\delta_u$) is a vector of the Euclidean space $\mathbb{R}^n$ (resp. $\mathbb{R}^m$). More precisely, it is the vector tangent to a piecewise smooth curve connecting a pair of points in $\mathbb{R}^n$ (resp. $\mathbb{R}^m$). The matrix $A\in\mathbb{R}^{n\times n}$ has components given, for every $(x,u)\in\mathbb{R}^n\times\mathbb{R}^m$, by 
\begin{equation*}
	A_{jk}(x,u)=\dfrac{\partial[f_j+b_ju_j]}{\partial x_k}(x,u)	
\end{equation*}
for indexes $j,k\in\mathbb{N}_{[1,n]}$.

\mypara The resulting system composed of Equations \eqref{eq:general system} and \eqref{eq:general system:differential} is analyzed on the state space spanned by the vector $(x,\delta_x)\in\mathbb{R}^n\times\mathbb{R}^n$.

\mypara Similarly to \eqref{eq:general system}, given a control $\delta_u$ for system \eqref{eq:general system:differential}, the solution to \eqref{eq:general system:differential} computed at time $t\geq0$, $(x,u)\in\mathbb{R}^n\times\mathbb{R}^n$ and issuing from the  initial condition $\delta_x\in\mathbb{R}^n$ is denoted by $\Delta_x(t,x,\delta_x,u,\delta_u)$.

\mypara Lyapunov stability notions of solutions to \eqref{eq:general system:differential} are similar to those of linear time-varying systems (LTVS) (see \cite{Hespanha:2009} for more information on LTVS).

\mypara The importance of the stability of \eqref{eq:general system:differential} for system \eqref{eq:general system} can be understood as follows. Given fixed controls $u$ and $\delta u$ for systems \eqref{eq:general system} and \eqref{eq:general system:differential}, respectively. If the norm of every solution $\Delta_x(t,x,\delta_x,u,\delta_u)$ converges to zero exponentially, as the time tends to infinity, then every pair of solutions to \eqref{eq:general system} converge to each other exponentially. The interested reader may address \cite{Lohmiller1998,Sontag2010} and references therein for further details.

\mypara A sufficient condition for the stability of \eqref{eq:general system:differential} is provided by analyzing the derivative of a particular function along the solutions of systems \eqref{eq:general system} and \eqref{eq:general system:differential}. This function is recalled from \cite{Forni2014} and \cite{Manchester2014a}.

@@

\mypara The existence a contraction metric for system \eqref{eq:general system} implies that every two solutions of this system converge to each other exponentially. The proof of this claim, for the case in which $u\equiv0$ and $\delta_u\equiv0$, can be found in \cite[Theorem 1]{Lewis1951}, and \cite[Theorems 5.7 and 5.33]{Reich2005}, and \cite[Lemma 3.3]{Isac2008}.

\mypara For the class of systems considered in this paper, the following kind of metric is of interest, since it also allows the design a feedback law for system \eqref{eq:general system:differential}.
